---
title: "WTP Boostrap"
author: "Rishi"
date: "01/30/2024"
output: html_document
---

# Initialize the libraries

```{r}

library(readxl)
library(ggplot2)
setwd('~/Documents/MSBA/Subjects/Winter/Advanced Statistics/Homework /Homework 2/')

```

# Read the data file

```{r}

# Read the design matrix
design_matrix <- readxl::read_excel("Preferences_Rishi_1.0.xlsx")
View(design_matrix)
```

# Run linear regression for conjoint analysis to compute pathworths

```{r}

# Initial conjoint model
conjoint_model <- lm(rank ~ . - `Profile Nos` - `Profile Specs` - category - score, data = design_matrix)
initial_summary <- summary(conjoint_model)
initial_partworths <- coef(conjoint_model)

```

# Calculate WTP for computed pathworths

```{r}

# Calculate price per utility
price_savings <- 2500 - 2000
price_partworth <- abs(initial_partworths[6])
price_per_utility <- price_savings / price_partworth

# Initial WTP calculations
initial_WTP <- abs(initial_partworths[2:6]) * price_per_utility

```

# Set the resampling limit and initialise the matrix for both Residual Bootstrap and Data bBootstrap

```{r}

# Bootstrap settings
bb <- 1000
n <- nrow(design_matrix)

# Initialize matrices for bootstrap results
bootstrap_WTPs_resid <- matrix(NA, nrow = bb, ncol = length(initial_WTP))
bootstrap_WTPs_data <- matrix(NA, nrow = bb, ncol = length(initial_WTP))

# Residuals for residual bootstrap
residuals <- conjoint_model$residuals
fitted_values <- conjoint_model$fitted.values

```

# Compute Residual Bootstrap 1000 times to get 95% CI for WTPs

```{r}

# Run the loop 1000 times
for(i in 1:bb) {
  # Residual Bootstrap
  new_rank_resid <- fitted_values + sample(residuals, n, replace = TRUE)
  model_resid <- lm(new_rank_resid ~ . - `Profile Nos` - `Profile Specs` - category - score, data = design_matrix)
  partworths_resid <- coef(model_resid)
  bootstrap_WTPs_resid[i, ] <- abs(partworths_resid[2:6]) * price_per_utility
  
  # Data Bootstrap
  sample_indices <- sample(1:n, n, replace = TRUE)
  bootstrap_sample <- design_matrix[sample_indices, ]
  model_data <- lm(rank ~ . - `Profile Nos` - `Profile Specs` - category - score, data = bootstrap_sample)
  partworths_data <- coef(model_data)
  bootstrap_WTPs_data[i, ] <- abs(partworths_data[2:6]) * price_per_utility
}

# Compute 95% CI for WTPs from both approaches
WTP_CI_resid_lower <- apply(bootstrap_WTPs_resid, 2, function(x) quantile(x, probs = 0.025, na.rm = TRUE))
WTP_CI_resid_upper <- apply(bootstrap_WTPs_resid, 2, function(x) quantile(x, probs = 0.975, na.rm = TRUE))
WTP_CI_data_lower <- apply(bootstrap_WTPs_data, 2, function(x) quantile(x, probs = 0.025, na.rm = TRUE))
WTP_CI_data_upper <- apply(bootstrap_WTPs_data, 2, function(x) quantile(x, probs = 0.975, na.rm = TRUE))
WTP_resid_mean <- apply(bootstrap_WTPs_resid,2, function(x) mean(x))
WTP_data_mean <- apply(bootstrap_WTPs_data,2, function(x) mean(x))

# Combine lower and upper CIs
WTP_CI_resid <- rbind(WTP_CI_resid_lower, WTP_CI_resid_upper)
WTP_CI_data <- rbind(WTP_CI_data_lower, WTP_CI_data_upper)

```

# Display the computed CIs

```{r}

# Convert the results in a dataframe for better readability
WTP_CI_resid_df <- data.frame(
  Attribute = names(initial_WTP),
  Lower_CI_Residual = WTP_CI_resid[1, ],
  Upper_CI_Residual = WTP_CI_resid[2, ])

WTP_CI_data_df <- data.frame(
  Attribute = names(initial_WTP),
  Lower_CI_Data = WTP_CI_data[1, ],
  Upper_CI_Data = WTP_CI_data[2, ])

WTP_CI_resid_df$Mean <- WTP_resid_mean
WTP_CI_data_df$Mean <- WTP_data_mean

# Print results using cat for clearer, more formatted output
cat("Residual Bootstrap WTP Confidence Intervals (Lower, Upper):\n")
print(WTP_CI_resid_df, row.names = FALSE)
cat("\nData Bootstrap WTP Confidence Intervals (Lower, Upper):\n")
print(WTP_CI_data_df, row.names = FALSE)

```

# Plotting the CIs for better understanding

```{r}

# Add a method column to each data frame
WTP_CI_resid_df$Method <- "Residual"
WTP_CI_data_df$Method <- "Data"

# Match column names 
colnames(WTP_CI_resid_df) <- colnames(WTP_CI_data_df) <- c("Attribute", "Lower_CI", "Upper_CI","Mean" , "Method")

# Combine the data frames
WTP_CI_combined_df <- rbind(WTP_CI_resid_df, WTP_CI_data_df)

# Plotting
ggplot(WTP_CI_combined_df, aes(x = Attribute, y = Mean, ymin = Lower_CI, ymax = Upper_CI, color = Method)) +
  geom_pointrange() +
  theme_minimal() +
  labs(title = "Bootstrap Confidence Intervals for WTP Estimates",
       x = "Attribute",
       y = "WTP Estimate (Mean of CI)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width = 0.2) +
  scale_color_manual(values = c("Residual" = "blue", "Data" = "red")) +
  facet_wrap(~Method, scales = "free_y")

```

# THE END